---
title: "Location_Prediction"
author: "Gregory Biland"
date: "2023-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set root directory to the parent of the R folder if needed, 
# or ensure the RProj sets the working directory correctly.
# knitr::opts_knit$set(root.dir = "..") # Uncomment if needed
```

## Packages

```{r include=FALSE}
# Data Manipulation
library(dplyr)
library(tidyverse)
library(data.table)
library(readr)
library(collapse)

# Spatial Data
library(sf)
library(spdep)
library(mapview)
library(tmap)
library(cartography)
library(ggmap)
library(ggspatial)
library(stars)
library(raster)
library(rasterVis)
library(spgwr)

# Modelling & ML
library(lme4)
library(caret)
library(ranger)
library(randomForest)
library(xgboost)
library(gbm)
library(glmnet)
library(nnet)
library(MLeval)
library(MLmetrics)
library(pROC)
library(boot)
library(effects)
library(sjPlot)

# Visualization
library(ggplot2)
library(gridExtra)
library(corrplot)
library(car)
library(ggsci)
library(viridis)
library(scico)
library(RColorBrewer)
library(ggthemes)  # Contains theme_map

# Parallel Processing
library(parallel)
library(future)
library(future.apply)
library(furrr)

# Utilities
library(xtable)
library(DHARMa)
library(classInt)
library(rlang)
```

## Setup & Data

```{r include=FALSE}
# Define file paths dynamically (Relative to the R script location)
# Assumes structure:
# Project/
#   R/
#     Location_Prediction_Clean.Rmd
#   Data/
base_path <- file.path("../Data")

if (!dir.exists(base_path)) {
  stop("Data directory not found. Please check working directory.")
}

folder_path_SHP <- file.path(base_path, "Accidents/SHP/")
folder_path_figures <- file.path(base_path, "Figures/Publication/")

# Ensure output directories exist
if(!dir.exists(folder_path_figures)) dir.create(folder_path_figures, recursive = TRUE)

# Load data
accident_data_recoded <- st_read(file.path(folder_path_SHP, "accident_data_recoded.shp"))
accident_data_recoded$Date <- as.character(accident_data_recoded$Date)

Acc_dist <- st_read(file.path(base_path,"Accidents/Acc_dist.shp"))
Acc_coords_df <- as.data.frame(st_coordinates(Acc_dist))

PAP <- st_read(file.path(base_path, "Accidents/PAP/pseudo_absence_points.shp"))

# Network & Topology
RoadNetwork <- st_read(file.path(base_path, "Topology/Roads_Car/RoadNetwork.shp"))
RedLight <- st_read(file.path(base_path, "Network/Traffic_lights/traffic_lights.shp"))
PedestrianCrossing <- st_read(file.path(base_path, "Network/Pedestrian_crossing/FGS.shp"))
StopSign <- st_read(file.path(base_path, "Network/Way_of_right/stop_signs.shp"))
Junction <- st_read(file.path(base_path, "Network/Way_of_right/intersections3050.shp"))
ROW <- st_read(file.path(base_path, "Network/Way_of_right/RVT.shp"))
TramTrack <- st_read(file.path(base_path, "Topology/Tram_lines/tram_rails.shp"))
TrafficArea <- st_read(file.path(base_path, "Network/Traffic_areas/Traffic_areas.shp"))
Exposure <- st_read(file.path(base_path, "Topology/GVM-ZH Unfälle/GVM18.shp"))
ComplexIntersection <- st_read(file.path(base_path, "Network/Roundabouts/Complex_Intersections.shp"))
BicycleNetwork <- st_read(file.path(base_path, "Topology/Cycling_network/bicycle_roads.shp"))
BusLane <- st_read(file.path(base_path, "Topology/Bus_Lines/Bus.shp"))
PPS <- st_read(file.path(base_path, "Network/Public parking spaces/PPS_RW.shp"))

# Boarders & Context
lake_ZH <- st_read(file.path(base_path, "Spatial Data/Boarders/See_ZH.shp")) %>% st_transform(crs = 2056)
stat_zonen <- st_read(file.path(base_path, "Spatial Data/Statistical Areas/statistical_areas.shp"))
city_borders <- st_read(file.path(base_path, "Spatial Data/Boarders/city_boarders.shp"))
zvv_stops <- st_read(file.path(base_path,"Network/ZVV_Haltestellen/Haltestellen_des_offentlichen_Verkehrs_-OGD.gpkg"))
one_way <- st_read(file.path(base_path, "Possibilities/oneway/one_way.gpkg"))
forest_ZH <- st_read(file.path(base_path, "Spatial Data/Boarders/Wald/Forrest_ZH.shp"))
buildings <- st_read(file.path(base_path, "Spatial Data/Boarders/Buildings/Buildings.shp"))
greenareas <- st_read(file.path(base_path, "Spatial Data/Boarders/Green_Area/greenarea.gpkg"))
buffer_accidents <- st_read(file.path(base_path, "Spatial Data/Boarders/Negative_Buffer/Buffer_Accidents_NGreen.shp"))
st_crs(forest_ZH) <- 2056
```

## Universal Functions

```{r}
# --- Helper Functions ---

# Join points to closes roads
join_points_to_roads <- function(points, roads, cols) {
  distances <- sf::st_distance(points, roads)
  closest_indices <- apply(distances, 1, function(x) {
    if (any(is.finite(x))) {
      which.min(x)
    } else {
      NA
    }
  })
  
  closest_roads <- roads[closest_indices, cols]
  joined_data <- cbind(points, closest_roads)
  return(joined_data)
}

# Filter probabilities
filter_probabilities <- function(df, threshold = 0.8) {
  df[df$predicted_accident_rate_gbm >= threshold, ]
}

# Normalize function
normalize <- function(x) {
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
}

# Calculate RMSE
calculate_rmse <- function(predictions, observations) {
  sqrt(mean((predictions - observations)^2, na.rm = TRUE))
}

# --- Plotting Functions ---

# Function to plot the map scenario
plot_scenario_map <- function(data_filtered, title_text, output_filename = NULL) {
  p <- ggplot() +
    geom_sf(data = city_borders, fill = alpha("white", .1), color = "#cccccc", size = 0.3) +
    geom_sf(data = forest_ZH, fill = alpha("green4", .2), color = NA) +
    geom_sf(data = lake_ZH, fill = alpha("deepskyblue", .3), color = NA) +
    geom_sf(data = RoadNetwork, color = "darkgrey", lwd = 0.2) +
    stat_density_2d(data = Acc_coords_df, aes(x = X, y = Y, fill = after_stat(level), alpha = after_stat(level)), 
                    geom = "polygon", h = c(500, 500)) +
    scale_fill_gradient(low = "yellow", high = "red") +
    scale_alpha_continuous(range = c(0, 1)) +
    geom_sf(data = data_filtered, size = 0.1, color = "black") +
    theme_map() +
    labs(subtitle = title_text) +
    theme(
      text = element_text(size = 12),
      plot.title = element_text(size = 15, hjust = 0.5, color = "black", face = "bold"),
      plot.subtitle = element_text(size = 13, hjust = 0.5, color = "black", face = "italic"),
      plot.caption = element_text(size = 7, hjust = 0),
      plot.title.position = "plot",
      plot.background = element_rect(fill = NA, color = "black"),
      panel.background = element_blank(),
      panel.grid = element_blank(),
      axis.text = element_blank(),
      axis.title = element_blank(),
      legend.position = "none"
    ) +
    coord_sf(datum = st_crs(2056)) +
    annotation_scale(location = "br", width_hint = 0.25, style = "ticks") +
    labs(caption = paste0("Source: Open Data Zurich, 2023\n", "ggplot2 ", packageVersion("ggplot2")))
  
  if (!is.null(output_filename)) {
    ggsave(file.path(folder_path_figures, output_filename), p, width = 8, height = 8)
  }
  return(p)
}

# Function to rasterize points
rasterize_scenario <- function(scenario, field = "predicted_accident_rate_gbm", fun = mean, res = 25) {
  example_points <- as(scenario, "Spatial")
  extent_grid <- st_bbox(scenario)
  
  example_raster <- raster(nrows = ceiling((extent_grid$ymax - extent_grid$ymin) / res),
                           ncols = ceiling((extent_grid$xmax - extent_grid$xmin) / res),
                           xmn = extent_grid$xmin, xmx = extent_grid$xmax,
                           ymn = extent_grid$ymin, ymx = extent_grid$ymax,
                           crs = crs(example_points))
  
  rasterized_grid <- rasterize(example_points, example_raster, field = field, fun = fun)
  return(rasterized_grid)
}

# Function to plot the grid map
plot_scenario_grid <- function(raster_obj, title_text, output_filename = NULL, fill_name = "Probability", palette = "lajolla", limits = c(0, 1)) {
  
  test_spdf <- as(raster_obj, "SpatialPixelsDataFrame")
  test_df <- as.data.frame(test_spdf)
  colnames(test_df) <- c("value", "x", "y")
  
  p <- ggplot() +
    geom_sf(data =  lake_ZH, fill = alpha("deepskyblue", .1), color = "#cccccc", size = 0.3) +
    geom_sf(data = city_borders, fill = NA, color = "darkgrey", size = 0.3) +
    geom_sf(data = RoadNetwork, color = "darkgrey", lwd = 0.2) +
    geom_sf(data = forest_ZH, fill = alpha("green4", .2), color = NA) +
    geom_tile(data = test_df, aes(x = x, y = y, fill = value), alpha = 0.8, color = NA) +
    scale_fill_scico(name = fill_name, limits = limits, palette = palette, direction = -1) +
    theme_minimal() +
    labs(subtitle = title_text) +
    theme(
      text = element_text(size = 12),
      plot.title = element_text(size = 15, hjust = 0.5, color = "black", face = "bold"),
      plot.subtitle = element_text(size = 13, hjust = 0.5, color = "black", face = "italic"),
      plot.caption = element_text(size = 7, hjust = 0),
      plot.title.position = "plot",
      plot.background = element_rect(fill = NA, color = "black"),
      panel.background = element_blank(),
      panel.grid = element_blank(),
      axis.text = element_blank(),
      axis.title = element_blank(),
      legend.title = element_text(size = 8)
    ) +
    coord_sf(datum = st_crs(2056)) +
    annotation_scale(location = "br", width_hint = 0.25, style = "ticks") +
    labs(caption = paste0("Source: Open Data Zurich, 2023\n", "ggplot2 ", packageVersion("ggplot2")))
  
  if (!is.null(output_filename)) {
    ggsave(file.path(folder_path_figures, output_filename), p, width = 8, height = 8)
  }
  return(p)
}

```

# MODEL_Preperation
## Adding Distances and Network/ Topology information:

```{r eval=FALSE, include=FALSE}
# Data preparation
# This chunk prepares the Pseudo Absence Points (PAP) and Accident Data (ACC) combined dataset.

ACC_5yr <- accident_data_recoded %>% filter(Year >= 2016)
PAP_ACC <- rbind(PAP, ACC_5yr)

# Initialize columns
cols <- c("RedLight", "PedestrianCrossing", "StopSign", "Junction", "ROW", "TramTrack", "ComplexIntersection", "BusLane", "PPS","PT_Stops")
PAP_ACC[cols] <- NA

# SF objects to calculate distances
sf_objects <- list(RedLight, PedestrianCrossing, StopSign, Junction, ROW, TramTrack, ComplexIntersection, BusLane, PPS,zvv_stops)

# Parallel processing setup
# Note: mclapply works on POSIX systems (Mac/Linux). For Windows, use parLapply.
n_cores <- detectCores() - 1
PAP_ACC_list <- mclapply(sf_objects, function(y) st_distance(x = PAP_ACC, y), mc.cores = n_cores)
PAP_ACC[cols] <- lapply(PAP_ACC_list, function(d) sapply(1:nrow(d), function(i) min(d[i,], na.rm = TRUE)))

# Spatial polygon processing
PAP_ACC <- st_join(PAP_ACC, TrafficArea, join = st_intersects)
PAP_ACC$zonentyp_t <- replace(PAP_ACC$zonentyp_t, is.na(PAP_ACC$zonentyp_t), 50)

# Join data to closest roads and other layers
cols_to_join <- c("SpedLmt", "RodWdth", "RodSrfc", "PPS")
PAP_ACC <- join_points_to_roads(PAP_ACC, RoadNetwork, cols_to_join)
PAP_ACC <- join_points_to_roads(PAP_ACC, Exposure, colnames(Exposure))
PAP_ACC <- join_points_to_roads(PAP_ACC, BicycleNetwork, "bkTyp_b")

# Create a 10m buffer around one-way streets
one_way_buffers <- st_buffer(one_way, dist = 10)

# Calculate the distance matrix between accidents and road lines
distances <- st_distance(PAP_ACC, one_way)

# Get the index of the closest road line for each accident point
closest_roads <- apply(distances, MARGIN = 1, which.min)

# Add the index of the closest road to the accidents data
PAP_ACC <- PAP_ACC %>%
  mutate(closest_road = closest_roads)

# Spatial join with one_way_buffers using the closest road index
PAP_ACC <- PAP_ACC %>%
  mutate(objectid = one_way_buffers$objectid[closest_road]) %>%
  mutate(direction = ifelse(st_intersects(geometry, one_way_buffers[closest_road,]) %>% lengths() > 0, 1, 2))

# Select columns for final output
selected_cols <- c("ID", "Date", "Year", "Month", "Week", "COVID", "accidnt", "qnr", "knr", 
                   "RedLight", "PedestrianCrossing", "StopSign", "Junction", "ROW", "TramTrack", 
                   "ComplexIntersection", "BusLane", "zonentyp_t", "PPS", "SpedLmt", "RodWdth", 
                   "RodSrfc", "bkTyp_b", "Exposition", "geometry", "accidnt","PT_Stops", "direction")
PAP_ACC_filtered <- PAP_ACC %>% dplyr::select(all_of(selected_cols))

# Write to file
output_path <- file.path(base_path, "Accidents/PAP_ACC.geojson")
st_write(PAP_ACC_filtered, output_path, delete_dsn = TRUE)

```

###### Accident Locations Correlation:

```{r}
# Load and subset data
# Load, subset, and preprocess data
PAP_ACC <- st_read(file.path(base_path, "Accidents/PAP_ACC.geojson")) %>%
  as.data.frame() %>%
  dplyr::select(-Date, -geometry, -Year, -Month, -Week) %>%
  filter(SpedLmt < 80, SpedLmt != 0) %>%
  mutate(across(where(is.numeric), 
                ~ (. - min(.)) / (max(.) - min(.)), 
                .names = "{.col}_scaled"),
         accidntFactor = as.factor(accidnt),
         accidntFactorClass = factor(accidntFactor, levels = c(0, 1), labels = c("Class0", "Class1")))

# Filter RoadNetwork data (ensure RoadNetwork is already loaded in setup)
RoadNetwork <- RoadNetwork %>%
  filter(SpedLmt < 80, SpedLmt != 0)

```

## LMER

```{r cars, eval=FALSE, include=FALSE}
# Create the generalized linear mixed model
lmer_1 <- lmer(accidnt ~ PedestrianCrossing_scaled + StopSign_scaled + TramTrack_scaled + ComplexIntersection_scaled +
  RodWdth_scaled + Exposition_scaled + zonentyp_t_scaled + ROW_scaled + PPS_scaled + bkTyp_b_scaled + BusLane_scaled + 
  SpedLmt_scaled + PT_Stops_scaled + direction + 
  RodWdth_scaled * SpedLmt_scaled + (SpedLmt_scaled | Exposition_scaled), 
  data = PAP_ACC, 
  control = lmerControl(optimizer = "Nelder_Mead"), 
  REML = TRUE)

# Print the model summary
summary(lmer_1)
AIC(lmer_1)

vif(lmer_1)
```

### Predictions Accident Locations

#### Distance Matrix & Random Points Calculation

```{r eval=FALSE, include=FALSE}
# Parameters
grid_sizes <- c(4, 5, 6, 8, 10, 15, 20, 30, 40, 50, 60, 70, 80)
point_counts <- c(15000, 30000, 45000, 100000, 200000) # Number of points
point_networks_random <- list()

for (grid_size in grid_sizes) {
  for (point_count in point_counts) {

    # Create a grid using the bounding box of the road buffer with the specified cell size
    grid <- st_make_grid(st_bbox(buffer_accidents), cellsize = c(grid_size, grid_size), square = TRUE)
    grid_sf <- st_sf(geometry = grid)
  
  # Intersect the grid with the road buffer
  # Note: ensure buffer_accidents and city_borders logic is correct
  grid_sf_clipped <- st_intersection(grid_sf, buffer_accidents) 

    if (nrow(grid_sf_clipped) >= point_count) {
      selected_cells <- grid_sf_clipped[sample(nrow(grid_sf_clipped), point_count), ]
    } else {
      warning("There are fewer than requested grid cells available after intersection. Using all available.")
      selected_cells <- grid_sf_clipped
    }
  
  # Create a random point within each selected cell
  generate_random_point <- function(cell) {
    bb <- st_bbox(cell)
    x <- runif(1, bb["xmin"], bb["xmax"])
    y <- runif(1, bb["ymin"], bb["ymax"])
    st_point(c(x, y))
  }
  
    points <- st_sfc(lapply(selected_cells$geometry, generate_random_point), crs = st_crs(selected_cells))
    point_network_random <- st_sf(geometry = points)
    
    # Assign to the list with a dynamic name
    point_networks_random[[paste0("point_network_random_", grid_size, "m_", point_count)]] <- point_network_random
  }
}
  
cols_dist <- c("RedLight", "PedestrianCrossing", "StopSign", "Junction", "ROW", "TramTrack", 
          "ComplexIntersection", "BusLane", "PPS", "PT_Stops")

cols_to_join <- c("SpedLmt", "RodWdth", "RodSrfc", "PPS")
sf_objects <- list(RedLight, PedestrianCrossing, StopSign, Junction, ROW, TramTrack, ComplexIntersection, BusLane, PPS, zvv_stops)

# Function to process a chunk of data
process_random_points_chunk <- function(data, sf_objects, cols_dist) {
  
  # Create a 10m buffer around one-way streets
  one_way_buffers <- st_buffer(one_way, dist = 10)

  # Calculate the distance matrix between accidents and road lines
  distances <- st_distance(data, one_way)

  # Get the index of the closest road line for each accident point
  closest_roads <- apply(distances, MARGIN = 1, which.min)

  # Add the index of the closest road to the accidents data
  data <- data %>%
    mutate(closest_road = closest_roads,
           objectid = one_way$objectid[closest_roads],
           direction = ifelse(st_intersects(geometry, one_way_buffers[closest_roads, ]) %>% lengths() > 0, 1, 2))
  
  # Calculate distances to features
  # Ensure plan() is called outside or here if not set
  # plan(multisession) # Use multisession for broader compatibility
  
  data_list <- future_lapply(sf_objects, function(y) st_distance(x = data, y = y))
  data[cols_dist] <- lapply(data_list, function(d) sapply(1:nrow(d), function(i) min(d[i,], na.rm = TRUE)))

  data <- st_join(data, TrafficArea, join = st_intersects)
  data$zonentyp_t <- replace(data$zonentyp_t, is.na(data$zonentyp_t), 50)

  data <- join_points_to_roads(data, RoadNetwork, cols_to_join)
  data <- join_points_to_roads(data, Exposure, colnames(Exposure))
  data <- join_points_to_roads(data, BicycleNetwork, "bkTyp_b")
  
  return(data)
}

# Set parallel plan once
plan(multisession, workers = max(1, detectCores() - 1))

# Main processing loop
for (grid_name in names(point_networks_random)) {
  
  data <- point_networks_random[[grid_name]]
  for (col in cols_dist) {
    data[[col]] <- NA
  }

  # Chunk processing
  chunk_size <- 1000
  num_chunks <- ceiling(nrow(data) / chunk_size)
  processed_data_list <- list()

  for (i in 1:num_chunks) {
    start_row <- ((i - 1) * chunk_size) + 1
    end_row <- min(i * chunk_size, nrow(data))

    data_chunk <- data[start_row:end_row, ]
    processed_data_list[[i]] <- process_random_points_chunk(data_chunk, sf_objects, cols_dist)
  }

  # Combine all processed chunks
  processed_data <- do.call(rbind, processed_data_list)

  # Write to file
  output_path <- file.path(base_path, "Prediction_Grid/Random", paste0(grid_name, ".geojson"))
  # Ensure directory exists
  dir.create(dirname(output_path), recursive = TRUE, showWarnings = FALSE)
  st_write(processed_data, output_path, delete_dsn = TRUE)
}

```

## Prediction evaluation:

```{r}
set.seed(123) # for reproducibility
trainIndex_AccPred <- createDataPartition(PAP_ACC$accidntFactorClass, p = 0.9, list = FALSE)
trainSet_AccPred <- PAP_ACC[trainIndex_AccPred, ]
testSet_AccPred <- PAP_ACC[-trainIndex_AccPred, ]

# Common Model Formula
model_formula <- accidntFactorClass ~ PedestrianCrossing + StopSign + TramTrack + ComplexIntersection +
              RodWdth + Exposition + zonentyp_t + ROW + PPS + bkTyp_b + BusLane + 
              SpedLmt + PT_Stops + direction +
              RodWdth * SpedLmt + SpedLmt * Exposition

# Evaluation Control
ctrl_eval <- trainControl(method="cv", 
                     summaryFunction=twoClassSummary, 
                     classProbs=TRUE,
                     savePredictions = TRUE)
```

#GBM

```{r}
paramGrid_gbm <- expand.grid(
  interaction.depth = 18,
  n.trees = 300,
  shrinkage = 0.05,
  n.minobsinnode = 20
)

# Train the model using caret
fit_GBM_eval <- train(model_formula, 
              data = trainSet_AccPred, 
              method = "gbm", 
              trControl = ctrl_eval,
              verbose = FALSE,
              tuneGrid = paramGrid_gbm,
              metric = "ROC")

# Predict on the test set
predictions_GBM <- predict(fit_GBM_eval, newdata = testSet_AccPred, type = "raw")
probabilities_GBM <- predict(fit_GBM_eval, newdata = testSet_AccPred, type = "prob")

# Performance
cm_GBM <- confusionMatrix(predictions_GBM, testSet_AccPred$accidntFactorClass)
print(cm_GBM)
```

#Random Forest

```{r}
paramGrid_rf <- expand.grid(mtry = 5)

# Train the model using caret
fit_RF_eval <- train(model_formula, 
              data = trainSet_AccPred, 
              method = "rf", 
              trControl = ctrl_eval,
              verbose = FALSE,
              tuneGrid = paramGrid_rf,
              metric = "ROC",
              ntree = 500)

predictions_RF <- predict(fit_RF_eval, newdata = testSet_AccPred, type = "raw")
probabilities_RF <- predict(fit_RF_eval, newdata = testSet_AccPred, type = "prob")

# Performance
cm_RF <- confusionMatrix(predictions_RF, testSet_AccPred$accidntFactorClass)
print(cm_RF)
```

#XGBoost

```{r}
paramGrid_xgb <- expand.grid(
  eta = 0.1,
  max_depth = 8,
  gamma = 5,
  colsample_bytree = 0.5,
  min_child_weight = 10,
  nrounds = 250,
  subsample = 0.5
)

# Train the model using caret
fit_XGB_eval <- train(model_formula, 
              data = trainSet_AccPred, 
              method = "xgbTree", 
              trControl = ctrl_eval,
              verbose = FALSE,
              tuneGrid = paramGrid_xgb,
              metric = "ROC")

predictions_XGB <- predict(fit_XGB_eval, newdata = testSet_AccPred, type = "raw")
probabilities_XGB <- predict(fit_XGB_eval, newdata = testSet_AccPred, type = "prob")

# Performance
cm_XGB <- confusionMatrix(predictions_XGB, testSet_AccPred$accidntFactorClass)
print(cm_XGB)
```

```{r}
## run MLeval
res <- evalm(list(fit_GBM_eval,fit_RF_eval,fit_XGB_eval), gnames=c('gbm','rf',"xgbTree"))

# ROC Curves
roc_curve_GBM <- roc(testSet_AccPred$accidntFactorClass, probabilities_GBM$Class1) 
roc_curve_RF <- roc(testSet_AccPred$accidntFactorClass, probabilities_RF$Class1) 
roc_curve_XGB <- roc(testSet_AccPred$accidntFactorClass, probabilities_XGB$Class1) 

# Convert ROC objects to data frames
df_GBM <- data.frame(tpr = roc_curve_GBM$sensitivities, fpr = roc_curve_GBM$specificities, Model = 'GBM')
df_RF <- data.frame(tpr = roc_curve_RF$sensitivities, fpr = roc_curve_RF$specificities, Model = 'RF')
df_XGB <- data.frame(tpr = roc_curve_XGB$sensitivities, fpr = roc_curve_XGB$specificities, Model = 'XGB')

roc_data <- rbind(df_GBM, df_RF, df_XGB)

# Create the ROC plot using ggplot2
ggplot(roc_data, aes(x = fpr, y = tpr, color = Model)) +
  geom_line() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "ROC Curve Comparison", x = "False Positive Rate", y = "True Positive Rate") +
  theme_minimal() +
  scale_color_manual(values = c("GBM" = "blue", "RF" = "red3", "XGB" = "lightgreen"))

```

## Prediction - Final Models

```{r}
ctrl_Pred <- trainControl(method="cv", 
                     summaryFunction=twoClassSummary, 
                     classProbs=TRUE,
                     savePredictions = TRUE)

# Train Final Models on Full Data
fit_GBM <- train(model_formula, 
              data = PAP_ACC, 
              method = "gbm", 
              trControl = ctrl_Pred,
              verbose = FALSE,
              tuneGrid = paramGrid_gbm,
              metric = "ROC")

fit_RF <- train(model_formula, 
              data = PAP_ACC, 
              method = "rf", 
              trControl = ctrl_Pred,
              verbose = FALSE,
              tuneGrid = paramGrid_rf,
              metric = "ROC",
              ntree = 500)

fit_XGB <- train(model_formula, 
              data = PAP_ACC, 
              method = "xgbTree", 
              trControl = ctrl_Pred,
              verbose = FALSE,
              tuneGrid = paramGrid_xgb,
              metric = "ROC")

```

### Prediction Processing Function

```{r}
# Generic function to process predictions for a scenario
process_scenario_prediction <- function(grid_size, point_count, modifier_function = NULL) {
  
  # Read the file
  fpath <- file.path(base_path, "Prediction_Grid/Random", paste0("point_network_random_", grid_size, "m_", point_count, ".geojson"))
  if(!file.exists(fpath)) {
     stop(paste("File not found:", fpath))
  }
  
  point_network <- st_read(fpath, quiet = TRUE) %>%
                   subset(SpedLmt < 80 & SpedLmt != 0)
  
  # Apply scenario modifications
  if (!is.null(modifier_function)) {
    point_network <- modifier_function(point_network)
  }
  
  # Models' predictions
  # Ensure fit_GBM, fit_RF, fit_XGB are available in the global environment
  point_network$predicted_accident_rate_gbm <- predict(fit_GBM, newdata = point_network, type = "prob")[,2]
  point_network$predicted_accident_rate_rf <- predict(fit_RF, newdata = point_network, type = "prob")[,2]
  point_network$predicted_accident_rate_xgb <- predict(fit_XGB, newdata = point_network, type = "prob")[,2]
  
  point_network <- st_set_geometry(point_network, "geometry")

  return(point_network)
}
```

```{r eval=FALSE, include=FALSE}
# Optional: run predictions for all grid-size / point-count combinations
grid_sizes <- c(4, 5, 6, 8, 10, 15, 20, 30, 40, 50, 60, 70, 80)
point_counts <- c(15000, 30000, 45000, 100000, 200000)
combos <- expand.grid(grid_size = grid_sizes, point_count = point_counts)

results_original_all <- lapply(seq_len(nrow(combos)), function(i) {
  gs <- combos$grid_size[i]
  pc <- combos$point_count[i]
  process_scenario_prediction(gs, pc)
})
names(results_original_all) <- paste0("grid_", combos$grid_size, "m_points_", combos$point_count)
```

### Predicting Original Grid

```{r}
# Original Scenario (No modification)
results_original <- process_scenario_prediction(8, 100000)
filtered_original <- filter_probabilities(results_original)

# Plot Map
scenario_original_map <- plot_scenario_map(filtered_original, "Original Scenario", "scenario_original.png")
print(scenario_original_map)

# Plot Grid
grid_original <- rasterize_scenario(results_original)
scenario_original_grid <- plot_scenario_grid(grid_original, "Original Scenario", "scenario_original_grid.png")
print(scenario_original_grid)
```

### RMSE
#### Random 

```{r eval=FALSE, include=FALSE}
# RMSE Calculation Chunk
# This section calculates RMSE for validation. Kept eval=FALSE as per original.

# Function to bootstrap RMSE
bootstrap_rmse <- function(observed, predicted) {
  # boot function requires a statistic function
  statistic_func <- function(data, indices) {
    d <- data[indices, ]
    return(calculate_rmse(d$predicted, d$observed))
  }
  
  set.seed(123)
  results <- boot(data = data.frame(observed = observed, predicted = predicted),
                  statistic = statistic_func,
                  R = 100)
  return(results)
}

# Wrapper to process grid and calculate RMSE
process_grid_rmse <- function(data_results, acc_dist, point_count_val) {
  
  data_results$ID <- seq_len(nrow(data_results))
  bbox <- st_bbox(data_results)
  
  # Make grid
  # Note: logic adapted to be more robust
  grid_sf_geom <- st_make_grid(st_as_sfc(bbox), cellsize = 25, square = TRUE)
  grid_sf <- st_sf(ID_GRID = seq_len(length(grid_sf_geom)), geometry = grid_sf_geom)
  
  # Count accidents per grid cell
  grid_with_accidents <- st_join(grid_sf, acc_dist, join = st_intersects)
  
  grid_with_accidents <- grid_with_accidents %>%
    group_by(ID_GRID) %>%
    summarise(accidents = ifelse(all(is.na(ID)), 0, 1), .groups = 'drop') # ID from Acc_dist
  
  # Map point results to grid cells to compare
  # This part assumes spatial relationship. 
  # If results_original are points, we join them to the grid to aggregate predictions per cell.
  
  points_in_grid <- st_join(data_results, grid_sf, join = st_intersects)
  
  # Aggregate predictions per grid cell (mean probability)
  grid_predictions <- points_in_grid %>%
    st_drop_geometry() %>%
    group_by(ID_GRID) %>%
    summarise(
      pred_gbm = mean(predicted_accident_rate_gbm, na.rm=TRUE),
      pred_rf = mean(predicted_accident_rate_rf, na.rm=TRUE),
      pred_xgb = mean(predicted_accident_rate_xgb, na.rm=TRUE),
      .groups = 'drop'
    )
  
  # Join accidents and predictions
  comparison <- merge(st_drop_geometry(grid_with_accidents), grid_predictions, by = "ID_GRID")
  comparison <- na.omit(comparison)
  
  rmse_gbm <- calculate_rmse(comparison$pred_gbm, comparison$accidents)
  rmse_rf <- calculate_rmse(comparison$pred_rf, comparison$accidents)
  rmse_xgb <- calculate_rmse(comparison$pred_xgb, comparison$accidents)
  
  return(data.frame(
    point_count = point_count_val,
    rmse_gbm = rmse_gbm,
    rmse_rf = rmse_rf,
    rmse_xgb = rmse_xgb
  ))
}

# Example usage
# res_rmse <- process_grid_rmse(results_original, Acc_dist, 200)
```


## Prediction Scenario

Scenarios: 
1. V30, Exposure 2000, Strasse 10% Breiter, Velotyp 3
2. V50, Exposure 8000, Strasse 20% Schmaler, Velotyp 1
3. V30
4.1. V30, Exposure 8000, Strasse 30% Breiter, Velotyp 1
4.2. V50, Exposure 2000, Strasse 20% Schmaler, Velotyp 2

### Scenario 1

Scenario where the speed is 30 in the entire city and the road width is 8m everywhere.

```{r}
# Modifier for Scenario 1
modify_scenario_1 <- function(data) {
  data$SpedLmt <- ifelse(data$SpedLmt >= 30, 30, data$SpedLmt)
  data$RodWdth <- data$RodWdth * 1.1
  data$Exposition <- 2000
  data$bkTyp_b <- 3
  return(data)
}

results_scenario_1 <- process_scenario_prediction(8, 100000, modify_scenario_1)
filtered_scenario_1 <- filter_probabilities(results_scenario_1)

# Plot
scenario_1_map <- plot_scenario_map(filtered_scenario_1, "Scenario 1", "scenario_1.png")
print(scenario_1_map)

scenario_1_grid <- plot_scenario_grid(rasterize_scenario(results_scenario_1), "Scenario 1", "scenario_1_grid.png")
print(scenario_1_grid)
```

### Scenario 2

```{r}
modify_scenario_2 <- function(data) {
  data$SpedLmt <- 50
  data$RodWdth <- data$RodWdth * 0.8
  data$Exposition <- ifelse(data$Exposition <= 8000, 8000, data$Exposition)
  data$bkTyp_b <- 1
  return(data)
}

results_scenario_2 <- process_scenario_prediction(8, 100000, modify_scenario_2)
filtered_scenario_2 <- filter_probabilities(results_scenario_2)

scenario_2_map <- plot_scenario_map(filtered_scenario_2, "Scenario 2", "scenario_2.png")
print(scenario_2_map)

scenario_2_grid <- plot_scenario_grid(rasterize_scenario(results_scenario_2, fun = max), "Scenario 2", "scenario_2_grid.png")
print(scenario_2_grid)
```


### Scenario 3

Scenario where the speed is 30 in the entire city.

```{r}
modify_scenario_3 <- function(data) {
  data$SpedLmt <- ifelse(data$SpedLmt >= 30, 30, data$SpedLmt)
  return(data)
}

results_scenario_3 <- process_scenario_prediction(8, 100000, modify_scenario_3)
filtered_scenario_3 <- filter_probabilities(results_scenario_3)

scenario_3_map <- plot_scenario_map(filtered_scenario_3, "Scenario 3", "scenario_3.png")
print(scenario_3_map)

scenario_3_grid <- plot_scenario_grid(rasterize_scenario(results_scenario_3), "Scenario 3", "scenario_3_grid.png")
print(scenario_3_grid)
```


### Scenario 4.1

```{r}
modify_scenario_4.1 <- function(data) {
  data$SpedLmt <- ifelse(data$SpedLmt >= 30, 30, data$SpedLmt)
  data$RodWdth <- data$RodWdth * 1.3
  data$Exposition <- 8000
  data$bkTyp_b <- 1
  return(data)
}

results_scenario_4.1 <- process_scenario_prediction(8, 100000, modify_scenario_4.1)
filtered_scenario_4.1 <- filter_probabilities(results_scenario_4.1)

scenario_4.1_map <- plot_scenario_map(filtered_scenario_4.1, "Scenario 4.1", "scenario_4.1.png")
print(scenario_4.1_map)

scenario_4.1_grid <- plot_scenario_grid(rasterize_scenario(results_scenario_4.1), "Scenario 4.1", "scenario_4.1_grid.png")
print(scenario_4.1_grid)
```

### Scenario 4.2

```{r}
modify_scenario_4.2 <- function(data) {
  data$SpedLmt <- 50
  data$RodWdth <- data$RodWdth * 0.8
  data$Exposition <- 2000
  data$bkTyp_b <- 2
  return(data)
}

results_scenario_4.2 <- process_scenario_prediction(8, 100000, modify_scenario_4.2)
filtered_scenario_4.2 <- filter_probabilities(results_scenario_4.2)

scenario_4.2_map <- plot_scenario_map(filtered_scenario_4.2, "Scenario 4.2", "scenario_4.2.png")
print(scenario_4.2_map)

scenario_4.2_grid <- plot_scenario_grid(rasterize_scenario(results_scenario_4.2), "Scenario 4.2", "scenario_4.2_grid.png")
print(scenario_4.2_grid)
```

--------------------------------------------------------------------------------------------------------------------------
# Comparison and Conclusion

## Differences in probabilities and distributions

```{r}
# Normalize function is already in global functions

# 1. Extract relevant probabilities
prob_original <- normalize(results_original$predicted_accident_rate_gbm)
prob_scenario_1 <- normalize(results_scenario_1$predicted_accident_rate_gbm)
prob_scenario_2 <- normalize(results_scenario_2$predicted_accident_rate_gbm)
prob_scenario_3 <- normalize(results_scenario_3$predicted_accident_rate_gbm)
prob_scenario_4.1 <- normalize(results_scenario_4.1$predicted_accident_rate_gbm)
prob_scenario_4.2 <- normalize(results_scenario_4.2$predicted_accident_rate_gbm)

probability_distribution <- ggplot() +
  geom_density(aes(x = prob_original, color = "Original")) +
  geom_density(aes(x = prob_scenario_1, color = "Scenario 1")) +
  geom_density(aes(x = prob_scenario_2, color = "Scenario 2")) +
  geom_density(aes(x = prob_scenario_3, color = "Scenario 3")) +
  geom_density(aes(x = prob_scenario_4.1, color = "Scenario 4.1")) +
  geom_density(aes(x = prob_scenario_4.2, color = "Scenario 4.2")) +
  geom_vline(xintercept = 0.8, linetype = "dashed", color = "black", linewidth = 0.5)+
  labs(subtitle = "Gradient Boosting (GB)",
       x = "Predicted Accident Probability",
       y = "Density", color = "Scenario") +
  scale_color_manual(values = c("Original" = "black", "Scenario 1" = "green3", "Scenario 2" = "red3", "Scenario 3" = "purple", "Scenario 4.1" = "orange1", "Scenario 4.2" = "orange3" )) +
  theme_minimal()+
      theme(
      text = element_text(size = 12),
      plot.title = element_text(size = 15, hjust = 0.5, color = "black", face = "bold"),
      plot.subtitle = element_text(size = 13, hjust = 0.5, color = "black", face = "italic"),
      plot.caption = element_text(size = 7, hjust = 0),
      plot.title.position = "plot",
      plot.background = element_rect(fill = NA, color = "black"),
        legend.position = "bottom",
        legend.key.size = unit(0.25, 'cm'),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm'),
        legend.title = element_text(size=8)) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))

print(probability_distribution)

ggsave(file.path(folder_path_figures, "probability_distribution.png"), probability_distribution, width = 8, height=4)

```

## Grid differences

```{r}
# Helper function for difference plots
plot_difference_grid <- function(diff_raster, title_text, filename = NULL) {
  difference_df <- as.data.frame(as(diff_raster, "SpatialPixelsDataFrame"))
  colnames(difference_df) <- c("value", "x", "y")
  
  p <- ggplot() +  
    geom_sf(data = RoadNetwork, color = "darkgrey", lwd = 0.2) +
    geom_tile(data = difference_df, aes(x = x, y = y, fill = value), color = NA) +
    scale_fill_scico(name = "Difference", limits = c(-1, 1), palette = "vik") +
    geom_sf(data = lake_ZH, fill = alpha("deepskyblue", .1), color = "#cccccc", size = 0.3) +
    geom_sf(data = city_borders, fill = NA, color = "darkgrey", size = 0.3) +
    geom_sf(data = forest_ZH, fill = alpha("green4", .2), color = NA) +
    theme_minimal()+
    labs(subtitle = title_text) +
    theme(
      text = element_text(size = 12),
      plot.title = element_text(size = 15, hjust = 0.5, color = "black", face = "bold"),
      plot.subtitle = element_text(size = 13, hjust = 0.5, color = "black", face = "italic"),
      plot.caption = element_text(size = 7, hjust = 0),
      plot.title.position = "plot",
      plot.background = element_rect(fill = NA, color = "black"),
      panel.background = element_blank(),
      panel.grid = element_blank(),
      axis.text = element_blank(),
      axis.title = element_blank(),
      legend.title = element_text(size = 8))+
    coord_sf(datum = st_crs(2056)) +
    annotation_scale(location = "br", width_hint = 0.25, style = "ticks")+
    labs(caption = paste0("Source: Open Data Zurich, 2023\n", "ggplot2 ", packageVersion("ggplot2")))
    
  print(p)
  if (!is.null(filename)) {
    ggsave(file.path(folder_path_figures, filename), p, width = 8, height=8)
  }
  return(p)
}

# Ensure we have rasterized versions. 
rasterized_scenario_original <- rasterize_scenario(results_original)
rasterized_scenario_1 <- rasterize_scenario(results_scenario_1)
rasterized_scenario_2 <- rasterize_scenario(results_scenario_2)
rasterized_scenario_3 <- rasterize_scenario(results_scenario_3)
```


### Best Worst- Best

```{r}
# Scenario 2 (Worst?) - Scenario 1 (Best?)
difference_BW <- rasterized_scenario_2 - rasterized_scenario_1
difference_grid_BW <- plot_difference_grid(difference_BW, "Scenario 1 to Scenario 2", "difference_grid_BW.png")
```

### Original- Best

```{r}
difference_OB <- rasterized_scenario_1 - rasterized_scenario_original
difference_grid_OB <- plot_difference_grid(difference_OB, "Original to Scenario 1", "difference_grid_OB.png")
```

### Original- Most Likely

```{r}
difference_OML <- rasterized_scenario_3 - rasterized_scenario_original
difference_grid_OML <- plot_difference_grid(difference_OML, "Original to Scenario 3", "difference_grid_OML.png")
```

### Best- Most Likely

```{r}
difference_BML <- rasterized_scenario_1 - rasterized_scenario_3
difference_grid_BML <- plot_difference_grid(difference_BML, "Scenario 3 to Scenario 1", "difference_grid_BML.png")
```


```{r}
# Composite Plots
plots_original_OML <- list(results_original$scenario_map, results_scenario_3$scenario_map)
plots_original_OW <- list(results_original$scenario_map, results_scenario_2$scenario_map)

## Point Distributions
comparison_OML <- grid.arrange(grobs = plots_original_OML, ncol = 2)
comparison_OW <- grid.arrange(grobs = plots_original_OW, ncol = 2)

## Grids
comparison_BW_grid <- grid.arrange(results_scenario_1$scenario_grid, results_scenario_2$scenario_grid, ncol = 2)
comparison_OB_grid <- grid.arrange(results_original$scenario_grid, results_scenario_1$scenario_grid, ncol = 2)
comparison_OML_grid <- grid.arrange(results_original$scenario_grid, results_scenario_3$scenario_grid, ncol = 2)

## Differences
difference_grids_OB_BW <- grid.arrange(difference_grid_OB, difference_grid_BW, ncol = 2)
difference_grids_OML_BML <- grid.arrange(difference_grid_OML, difference_grid_BML, ncol = 2)

# Save
ggsave(file.path(folder_path_figures, "comparison_OML.png"), comparison_OML, width = 8, height=8)
ggsave(file.path(folder_path_figures, "comparison_OW.png"), comparison_OW, width = 8, height=8)
ggsave(file.path(folder_path_figures, "comparison_BW_grid.png"), comparison_BW_grid, width = 8, height=8)
ggsave(file.path(folder_path_figures, "comparison_OB_grid.png"), comparison_OB_grid, width = 8, height=8)
ggsave(file.path(folder_path_figures, "comparison_OML_grid.png"), comparison_OML_grid, width = 8, height=8)
ggsave(file.path(folder_path_figures, "difference_grids_OB_BW.png"), difference_grids_OB_BW, width = 8, height=8)
ggsave(file.path(folder_path_figures, "difference_grids_OML_BML.png"), difference_grids_OML_BML, width = 8, height=8)

```
